---
title: 'Assignment 7: Stability'
author: "Colleen McCamy"
date: "2023-06-04"
output: html_document
---

```{r}

# loading libraries
library(tidyverse)
library(sensitivity)
library(tidyverse)
library(deSolve)
library(ggpubr)

```


```{r}

# sourcing in the function
path = "/Users/colleenmccamy/Documents/MEDS/classes/spring/eds-230-modeling/assignments/eds230-stability-modeling"
source(paste0(path, "/R/lotvmodK_hunt.R"))

```

```{r}

# setting parameters
rprey <- 0.95
alpha <- 0.01
eff <- 0.6
pmort <- 0.4
K <- 2000
pred_min <- 1000
hunt <- 0.5

# creating it to be a dataframe
pars = tibble(rprey=rprey, alpha=alpha, eff=eff, pmort=pmort, K=K, prey_min = prey_min, hunt = hunt)

# setting timescale
months = seq(from=1, to=300, by  = 1)


# establishing population at time 0
pop0 <- c(prey=2000, pred=200)

# integrating the model
res <- ode(func=lotvmodK_hunt, y=pop, times=months, parms=pars)

### --------------------- 
### sensitivity analysis 
### ---------------------

# sample 1
hunt_range <- runif(min = 0.01, max = 0.7, n = 200)
min_range <- runif(min = 1, max = 2000, n = 200)
X1 <-cbind.data.frame(hunt_range=hunt_range, min_range=min_range)

# sample 2
hunt_range <- runif(min = 0.01, max = 0.7, n = 200)
min_range <- runif(min = 1, max = 2000, n = 200)
X2 <-cbind.data.frame(hunt_range=hunt_range, min_range=min_range)

# creating sobol object
sensitivity_object <- sobolSalt(model = NULL,X1, X2, nboot = 300)

# selecting random values to test
values <- as.data.frame(sensitivity_object$X) |> 
  sample_n(20, replace = FALSE)
# Add column names
colnames(values) <- c("hunt", "min_pred")

# creating results shell
results <- list()

# iterating over the 20 random rows of the hunting and minimum pred values
for (i in 1:nrow(values)) {

  hunt <- values$hunt[i]
  pred_min <- values$min_pred[i]


  pars <- tibble(
    rprey = rprey,
    alpha = alpha,
    eff = eff,
    pmort = pmort,
    K = K,
    pred_min = pred_min,
    hunt = hunt
  )


  res <- ode(func = lotvmodK_hunt, y = pop0, times = months, parms = pars)


  key <- paste0("hunt_", hunt, "_preymin_", prey_min)
  results[[key]] <- as.data.frame(res)
}

# storing the results as a dataframe
results_df <- do.call(rbind, results)

```



Explore how different hunting levels and different minimum prey populations (before hunting is allowed) are likely to effect the stability of the populations of both predator and prey. Use this exploration to recommend a hunting target that will be sustainable (e.g leave you with a stable prey and predator population)

You can assume the following rprey=0.95, alpha=0.01, eff=0.6,pmort=0.4, K=2000,

A key challenge is how you might want to define stability? Its up to you but you will need to write a sentence to explain why you chose the measure that you did.  It could be something as simple as maintaining a population above some value 50 years into the future.

Submit the Rmarkdown that documents your exploration (e.g how you tested different hunting levels and how you defined a stability metric). At the end of your Rmarkdown present a recommendation for sustainable hunting and 1-2 sentences to justify your recommendation based on your analysis. 

