---
title: 'Assignment 7: Stability'
author: "Colleen McCamy"
date: "2023-06-04"
output: html_document
---

```{r, include=FALSE}

# loading libraries
library(tidyverse)
library(sensitivity)
library(tidyverse)
library(deSolve)
library(ggpubr)
library(ggplot2)

```


```{r, warning=FALSE}

# sourcing in the function
path = "/Users/colleenmccamy/Documents/MEDS/classes/spring/eds-230-modeling/assignments/eds230-stability-modeling"
source(paste0(path, "/R/lotvmodK_hunt.R"))

```

```{r, warning=FALSE}

### --------------------- 
### running the hunting model
### ---------------------

# setting parameters
rprey <- 0.95
alpha <- 0.01
eff <- 0.6
pmort <- 0.4
K <- 2000
pred_min <- 1000
hunt <- 0.5

# creating it to be a dataframe
pars = tibble(rprey=rprey, alpha=alpha, eff=eff, pmort=pmort, K=K, pred_min = pred_min, hunt = hunt)

# setting timescale
months = seq(from=1, to=300, by=1)

# establishing population at time 0
pop0 <- c(prey=2000, pred=200)

# integrating the model
res <- ode(func=lotvmodK_hunt, y=pop0, times=months, parms=pars)

### --------------------- 
### sensitivity analysis 
### ---------------------

# sample 1
hunt_range <- runif(min = 0.01, max = 0.7, n = 200)
min_range <- runif(min = 1, max = 2000, n = 200)
X1 <-cbind.data.frame(hunt_range=hunt_range, min_range=min_range)

# sample 2
hunt_range <- runif(min = 0.01, max = 0.7, n = 200)
min_range <- runif(min = 1, max = 2000, n = 200)
X2 <-cbind.data.frame(hunt_range=hunt_range, min_range=min_range)

# creating sobol object
sensitivity_object <- sobolSalt(model = NULL,X1, X2, nboot = 300)

# selecting random values to test
values <- as.data.frame(sensitivity_object$X) |> 
  sample_n(20, replace = FALSE)
# Add column names
colnames(values) <- c("hunt", "min_pred")

# creating results shell
results <- list()

# iterating over the 20 random rows of the hunting and minimum pred values
for (i in 1:nrow(values)) {

  hunt <- values$hunt[i]
  pred_min <- values$min_pred[i]


  pars <- tibble(
    rprey = rprey,
    alpha = alpha,
    eff = eff,
    pmort = pmort,
    K = K,
    pred_min = pred_min,
    hunt = hunt
  )


  res <- ode(func = lotvmodK_hunt, y = pop0, times = months, parms = pars)


  key <- paste0(hunt, ",", pred_min)
  results[[key]] <- as.data.frame(res)
}

# storing the results as a dataframe
results_df <- do.call(rbind, results)

# data wrangling
results_df <- results_df |> 
  rownames_to_column(var = "index")
results_df <- separate(results_df, index, into = c("hunt", "min_pred"), 
                       sep = ",", remove = FALSE)
results_df <- results_df |> 
  select(-index)

```

### Defining Stability
```{r, warning=FALSE}

# defining stability
stability_metric <- mapply(function(prey, pred) prey >= (3 * pred), results_df$prey, results_df$pred)

# applying stability metric
results_df <- results_df |> 
  mutate(stable = ifelse(stability_metric, "stable", "unstable"))

# stability averages
stable_avg <- results_df |> 
  filter(stable == "stable") |> 
  group_by(stable) |> 
  summarize(hunt = max(hunt),
            min_pred = min(min_pred))

```

### Discussion
Based on my metric of stability such that predators need to be at least 1/3 of the prey population for the population to be stable, I have found that in order to keep the population stable there is a maximum hunting rate of about 0.57 with a minimum population of predators at 1135 units of predators.

I would suggest that hunting limits and policies ensure that hunt rates do not exceed this maximum and that the minimum population does not dip below the minimum for a stable population of predators and prey.

